---
title: "Quickly visualize solutions"
output: pdf_document
date: "2023-07-26"
---

Take a glimpse at one of the solutions (reproduce Figure 1 panel B)!

```{r include = FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(sf)
library(raster)
library(fasterize)
library(terra)
require(viridisLite)
library(MetBrewer)
library(ggridges)
library(rasterVis)
library(colorspace)
library(scico)
```

Read in one of the solution

```{r}
sol <- "../data/solutions/sol/SI-sol/sol_carbon_0.5_restoration_0.01_production_NUTS2adj_country_FLEX_wetlands_TRUE_onlyrestoration_FALSE_scenario_Baseline_norest_highs_f455.csv"

#sol <- list.files("../data-formatted/sol/", pattern = "globiomICflat_gurobi_f455.csv", full.names = T)
solution <- read_csv(sol) |>
  dplyr::select(id, solution_1_z1:solution_1_z26)

sum(solution$solution_1_z13)


ic <- read_csv("../data/outputs/2-zones/PU_lc_intensity.csv") |> rename(pu = PUID)


pu_in_EU <- read_csv("../data/formatted-data/pu_in_EU.csv") 

ic <- ic |>
  left_join(pu_in_EU) |>
  rename(id = EU_id) |>
  dplyr::select(-c(pu, nuts2id)) |>
  drop_na(id)

zone_id <- read_csv("../data/formatted-data/zone_id.csv") 
zone_id$zone
colnames(solution) <- c("id", (zone_id$zone))
```



```{r include = FALSE, cache = TRUE}
MAES <- readRDS("../data/EEA_habitatpreferences.rds") |>
  filter(typeasso %in% c("Preferred", "Suitable")) |>
  mutate(maes_label = recode(maes_label, 
                             `Sparserly vegetated land` = "SparseVeg",
                             `Heathland and shrub` = "HeathlandShrub",
                             `Rivers and lakes` = "RiversLakes",
                             `Woodland and forest` = "WoodlandForest",
                             `Rivers and lakes` = "RiversLakes",
                             `Marine inlets and transitional waters` = "MarineTransitional"))
## fix naming of ../data/SpeciesData/BirdsDirectiveSpeciesThreats.csv## fix naming of species to match above
MAES$speciesname <- gsub(" ", "_", MAES$speciesname, fixed=TRUE)

PU_plot <- raster("../data/formatted-data/PU_raster.tif")
PU_template <- rasterToPolygons(PU_plot) |>
  st_as_sf() |>
  rename(id = PU_raster)
pu_in_EU <- read_csv("../data/formatted-data/pu_in_EU.csv")

colors <- c("grey", met.brewer(name="Isfahan1",type="continuous"))
```

```{r fig.width=10, fig.height=10}
solution_table <- pu_in_EU |> rename(id = pu) |> #plot_data |> 
  left_join(PU_template) |> 
  dplyr::select(-id) |>
  rename(id = EU_id) |>
  left_join(solution) |> 
  pivot_longer(-c(nuts2id:geometry)) 

solution_table_plot <- solution_table |> 
  mutate(zone = name) |>
  separate(name, c('maes_label', 'intensity', 'action'), sep = "_") |>
  mutate(maes_label = ifelse(maes_label %in% c("RiversLakes", "MarineTransitional"), "RiversLakesMarine", maes_label)) |>
  mutate(zone = paste0(maes_label,"_", intensity,"_", action)) |>
  group_by(id, maes_label, intensity, action) |> mutate(value = value) |> ungroup() |>
  unique()
```

Quick bar plot of zones

```{r width = 10}
as_tibble(solution_table_plot) |>
  group_by(action) |>
  summarise(area = sum(value)/41046)

bar_plot <- as_tibble(solution_table_plot) |>
  group_by(zone, maes_label, intensity, action) |>
  summarise(area = sum(value)) |>
  mutate(action = ifelse(action == "lockin", "production", action)) |>
  ungroup() |> filter(area >0) |> 
  mutate(intensity = ifelse(intensity == "primary", "natural", intensity)) |>
  mutate(name = paste0(maes_label, " \n (", intensity, ")")) |>
  ggplot(aes(x = reorder(name, -area), y = area/10, fill = action)) + geom_bar(stat="identity", width = 0.7) +
  theme_classic() + 
  labs(x = element_blank(), y = "area (1000 km^2)") +
  theme(legend.position = c(0.8,0.8),
        legend.title = element_blank()) +
    scale_fill_manual(values = c(met.brewer(name="Kandinsky",n=4,type="continuous"))[c(1,3,2)]) + coord_flip()

bar_plot
```

```{r}
bar_plot <- as_tibble(solution_table_plot) |>
  group_by(action) |>
  summarise(area = sum(value)) |>
  mutate(action = ifelse(action == "lockin", "urban", action)) |>
  ungroup() |> filter(area >0) |> mutate(tot_area = sum(area)) |>
  mutate(perc = area/tot_area*100) |>
  #mutate(intensity = ifelse(intensity == "primary", "natural", intensity)) |>
  #mutate(name = paste0(maes_label, " \n (", intensity, ")")) |>
  ggplot(aes(x = reorder(action, -perc), y = perc)) + geom_bar(stat="identity", width = 0.7) +
  theme_classic() + 
  labs(x = element_blank(), y = "percent area") +
  theme(legend.position = c(0.8,0.8),
        legend.title = element_blank()) 

bar_plot
```

Maps of solution

```{r}
solution_raster <- fasterize(st_as_sf(solution_table_plot), PU_plot, field = "value", by = "zone")
solution_raster <- rast(solution_raster)
names(solution_raster)

restore <- c(1:10)
produce <- c(12:18,25)
conserve <- c(19:24)
```

Color stuff
```{r}
colors <- c("grey", met.brewer(name="VanGogh3",n=20,type="continuous"))
myPal <- colors
myTheme <- rasterTheme(region = myPal)
```


```{r}
restore <- solution_raster[[restore]]
conserve <- solution_raster[[conserve]]
produce <- solution_raster[[produce]]
```

Just a check to make sure the solutions are summing to 1...

```{r}
solu_sum <- app(rast(solution_raster), 'sum')
plot(solu_sum)
```

Check out restoration zones and overall restoration

```{r}
library(colorspace)
library(scico)
cols.v <- c("grey",scico(20, direction = -1,palette = "batlow"))
#colors <- c("grey", met.brewer(name="Isfahan1",n=20,type="continuous"))

colors <- c("grey", met.brewer(name="VanGogh3",n=20,type="continuous"))
myPal <- colors
myTheme <- rasterTheme(region = myPal)

p1 <- levelplot(restore, par.settings = myTheme,xlab=NULL, ylab=NULL, scales=list(draw=FALSE))

p2 <- levelplot(conserve, par.settings = myTheme,xlab=NULL, ylab=NULL, scales=list(draw=FALSE))

p3 <- levelplot(produce, par.settings = myTheme,xlab=NULL, ylab=NULL, scales=list(draw=FALSE))

solu_sum <- app(solution_raster, 'sum')
plot(solu_sum)
```

```{r restzones, fig.width=10, fig.height = 5, dpi = 400}
p1
```

```{r rest_ov, fig.width=5, fig.height = 5.5, dpi = 400}
rest_agg <- app(restore, sum)
plot(rest_agg, col = colors)
```

Check out conservation zones and overall conservation

```{r consvzones, fig.width=10, fig.height = 5, dpi = 400}
levelplot(conserve, par.settings = myTheme,xlab=NULL, ylab=NULL, scales=list(draw=FALSE))
```

```{r cons_ov, fig.width=5, fig.height = 5.5, dpi = 400}
conserve_agg <- app(conserve, sum)
plot(conserve_agg, col = colors)
```

Check out production zones and overall production

```{r prodzones, fig.width=10, fig.height = 5, dpi = 400}
levelplot(produce, par.settings = myTheme,xlab=NULL, ylab=NULL, scales=list(draw=FALSE))
```

```{r prod_ov, fig.width=5, fig.height = 5.5, dpi = 400}
prod_agg <- app(produce, sum)
plot(prod_agg, col = colors)
```

reshuffling
```{r fig.width=6, fig.height = 6, dpi = 400}
solution_rs <- solution |>
  pivot_longer(-id) |>
  rename(value_sol = value) |>
  separate(name, c('maes_label', 'intensity', 'action'), sep = "_") |>
  mutate(maes_label = ifelse(maes_label %in% c("RiversLakes", "MarineTransitional"), "RiversLakesMarine", maes_label)) |>
  #mutate(zone = paste0(maes_label,"_", intensity,"_", action)) |>
  mutate(name = paste0(maes_label,"_", intensity))

solution_rs <- solution_rs |>
  group_by(id, name) |>
  summarise(value_sol = sum(value_sol, na.rm = T)) |> 
  ungroup()

unique(solution_rs$name)

ic_rs <- ic |> dplyr::select(-Status) |>
  pivot_longer(-id) |>
  rename(value_ic = value) |>
  mutate(name = ifelse(name %in% c("MarineTransitional_natural", "RiversLakes_natural"), "RiversLakesMarine_natural", name)) |>
  full_join(solution_rs) |> mutate(value_ic = replace_na(value_ic,0),
                value_sol= replace_na(value_sol, 0),
                diff = value_sol- value_ic) |> dplyr::select(-c(value_sol,value_ic))

ic_rs_wide <- ic_rs |> pivot_wider(names_from = name, values_from = diff)
ic_rs |>
  ggplot(aes(x = diff)) + geom_histogram(bins=50) + facet_wrap(~name) +
  labs(y = "number of pixels", x = "solution value - initial conditions") +
  theme_classic()

```



calculate shortfall

```{r}
targets <- read_csv("../data/formatted-data/targets_split_formatted.csv")
pu_in_EU <- read_csv("../data/formatted-data/pu_in_EU.csv")
raster <- raster("../data/landcover/10km/Corine_2018_cropland.tif")

pu <- read_fst("../data/formatted-data/pu_data.fst") |>
  left_join(pu_in_EU) |>
  rename(id = EU_id) |>
  dplyr::select(-c(pu, nuts2id)) |>
  drop_na(id)

rij <- read_fst("../data/formatted-data/features_split.fst") |>
  rename(species = feature) |>
  #mutate(amount = round(amount)) |>
  left_join(pu_in_EU) |>
  mutate(pu = EU_id) |>
  dplyr::select(pu, species, zone, amount) |>
  drop_na(pu) |>
  mutate(amount = replace_na(amount, 0))

rij |> filter(species == 999999) |> ggplot(aes(x = amount)) + geom_histogram() + facet_wrap(~zone)
## Calculate representation and save

filelist_temp <-  "../data/solutions/sol/SI-sol/sol_carbon_0.5_restoration_0.01_production_NUTS2adj_country_FLEX_wetlands_TRUE_onlyrestoration_FALSE_scenario_Baseline_norest_highs_f455.csv"

shortfalls <- list()

#  join targets and spp info
for (i in 1:length(filelist_temp)) {
  filenm <- unlist(strsplit(filelist_temp[[i]], "_"))
  carbon_weight <- filenm[[3]]
  country_contraint <- filenm[[9]]
  scenario <- filenm[[15]]
  future <- str_sub(filenm[[18]], 1,-5)

  solution <- read_csv(filelist_temp[[i]])

  solution <- solution |>
    rowid_to_column("pu") |>
    dplyr::select(pu, solution_1_z1:solution_1_z26) |>
    pivot_longer(-pu) |>
    mutate(zone = str_sub(name, 13)) |>
    dplyr::select(-name) |>
    mutate(zone = as.numeric(zone))

  rij_sol <- rij |>
    left_join(solution, by = c("pu", "zone"))

  contribution <- rij_sol |>
    mutate(contribution = amount*value) |>
    group_by(species) |>
    summarise(rep = sum(contribution))|>
    left_join(targets) |>
    mutate(shortfall = rep-target) |>
    mutate(shortfall_perc = shortfall/target) |>
    mutate(carbon_weight = carbon_weight,
           scenario = scenario,
           country_contraint = country_contraint,
           future = future) |>
    mutate(target_met = ifelse(shortfall<0,"no", "yes")) |>
    mutate(spp = ifelse(species != 999999, "biodiv", "carbon"))

  shortfalls[[i]] <- contribution
}

shortfalls <- bind_rows(shortfalls, .id = "column_label")
rep_IC <- read_csv("../data/solutions/representation_IC.csv")
rep <- shortfalls

```


```{r}
######################### set up files ##############
zone_id <- read_csv("../data/formatted-data/zone_id.csv")
spp_lookup_count <- read_csv("../data/plotting/spp_lookup.csv") |>
  filter(typeasso %in% c("Preferred", "Suitable"))|>
  group_by(speciesname) |> count() |>
  ungroup() |>
  mutate(n = ifelse(n>0,1,0))

spp_lookup <- read_csv("../data/plotting/spp_lookup.csv") |>
  dplyr::select(speciesname, speciesgroup, taxon_id) |>
  left_join(spp_lookup_count) |> unique() #|>
  #pivot_wider(names_from = maes_label, values_from = n)

BR_lookup <- read_csv("../data/plotting/bioregion_code.csv") |>
  rename(bioregion = BRIDnum) |>
  dplyr::select(code, bioregion)
country_lookup <- read_csv("../data/plotting/country_code.csv")
# plotting_data <- read_csv("data/plotting/plotting_data_ref_f455.csv")

PU_template <- raster("../data/landcover/10km/Corine_2018_cropland.tif") |>
  rasterToPolygons() |>
  st_as_sf() |>
  #st_transform(crs = st_crs(natura)) |>
  dplyr::mutate(id = seq(1:length(geometry))) |>
  dplyr::select(-Corine_2018_cropland)
raster <- raster("../data/landcover/10km/Corine_2018_cropland.tif")

pu_in_EU <- read_csv("../data/formatted-data/pu_in_EU.csv")

PU_plot <- pu_in_EU |> rename(id = pu) |> #plot_data |>
  left_join(PU_template) |>
  dplyr::select(EU_id, geometry) |> st_as_sf()

PU_plot <- fasterize(st_as_sf(PU_plot), raster, field = "EU_id")

bioregion <-read_csv("../data/formatted-data/pu_in_EU_BR.csv") |>
  dplyr::select(pu, BR_ID) |>
  rename(id = pu)

country <- read_csv("../data/formatted-data/linear_constraints/pu_pasture_low_budget_data.csv") |>
  dplyr::select(NUTS_ID, pu) |>
  mutate(country = substr(NUTS_ID, start = 1, stop = 2))|>
  rename(id = pu) |> dplyr::select(-NUTS_ID)

PU_template_EU <- pu_in_EU |>
  rename(id = pu) |> #plot_data |>
  left_join(PU_template) |>
  dplyr::select(-id) |>
  rename(id = EU_id)

colors <- c("grey", met.brewer(name="Isfahan1",type="continuous"))

rep_IC <- read_csv("../data/solutions/representation_IC.csv")

rep_IC_formatted <- rep_IC |>
  dplyr::select(-c(rep, target, shortfall)) |>
  filter(species != 999999) |>
  mutate(taxon_id = ifelse(species > 10000000000,
                           substr(species, start = 1, stop = 7),
                           substr(species, start = 1, stop = 6)),
         bioregion = ifelse(species > 10000000000,
                            substr(species, start = 8, stop = 9),
                            substr(species, start = 7, stop = 8)),
         country_code = ifelse(species > 10000000000,
                               substr(species, start = 10, stop = 11),
                               substr(species, start = 9, stop = 10))) |>
  mutate(taxon_id = as.numeric(taxon_id),
         bioregion = as.numeric(bioregion),
         country_code = as.numeric(country_code)) |>
  left_join(BR_lookup, multiple = "all") |>
  left_join(spp_lookup, multiple = "all")

rep_formatted <- rep |>
  dplyr::select(-c(column_label, rep, target, shortfall)) |>
  filter(species != 999999) |>
  mutate(taxon_id = ifelse(species > 10000000000,
                           substr(species, start = 1, stop = 7),
                           substr(species, start = 1, stop = 6)),
         bioregion = ifelse(species > 10000000000,
                            substr(species, start = 8, stop = 9),
                            substr(species, start = 7, stop = 8)),
         country_code = ifelse(species > 10000000000,
                               substr(species, start = 10, stop = 11),
                               substr(species, start = 9, stop = 10))) |>
  mutate(taxon_id = as.numeric(taxon_id),
         bioregion = as.numeric(bioregion),
         country_code = as.numeric(country_code)) |>
  left_join(BR_lookup, multiple = "all") |>
  left_join(spp_lookup, multiple = "all")
```

```{r}
##################### abstract statistics ##############################
ic_join <- rep_IC |>
  rename(rep_ic = rep,
         shortfall_ic = shortfall,
         shortfall_perc_ic = shortfall_perc) |>
  dplyr::select(species, rep_ic, shortfall_ic, shortfall_perc_ic)

## carbon percentage change
c_perc <- rep |>
  left_join(ic_join) |>
  filter(spp == "carbon") |>
  filter(rep_ic > 0) |>
  mutate(rep_diff = (rep - rep_ic)/rep_ic) |>
  group_by(spp, scenario, carbon_weight, country_contraint, future) |>
  summarise(rep_diff = rep_diff*100) |>
  arrange(-rep_diff)

## conservation status
b_status <- rep |>
  left_join(ic_join) |>
  filter(species != 999999) |>
  mutate(target_met_ic = ifelse(shortfall_perc_ic < 0, 0, 1),
         target_met = ifelse(shortfall_perc < 0, 0, 1)) |>
  filter(target_met_ic<1) |>
  mutate(improved = (target_met-target_met_ic)) |>
  group_by(spp, scenario, carbon_weight, country_contraint, future) |>
  summarise(improved = sum(improved)/length(unique(species))) |>
  arrange(-improved)
```


```{r}
rep_IC_formatted <- rep_IC |>
  dplyr::select(-c(rep, target, shortfall)) |>
  filter(species != 999999) |>
  mutate(taxon_id = ifelse(species > 10000000000,
                           substr(species, start = 1, stop = 7),
                           substr(species, start = 1, stop = 6)),
         bioregion = ifelse(species > 10000000000,
                            substr(species, start = 8, stop = 9),
                            substr(species, start = 7, stop = 8)),
         country_code = ifelse(species > 10000000000,
                               substr(species, start = 10, stop = 11),
                               substr(species, start = 9, stop = 10))) |>
  mutate(taxon_id = as.numeric(taxon_id),
         bioregion = as.numeric(bioregion),
         country_code = as.numeric(country_code)) |>
  left_join(BR_lookup, multiple = "all") |>
  left_join(spp_lookup, multiple = "all")

rep_formatted <- rep |>
  dplyr::select(-c(column_label, rep, target, shortfall)) |>
  filter(species != 999999) |>
  mutate(taxon_id = ifelse(species > 10000000000,
                           substr(species, start = 1, stop = 7),
                           substr(species, start = 1, stop = 6)),
         bioregion = ifelse(species > 10000000000,
                            substr(species, start = 8, stop = 9),
                            substr(species, start = 7, stop = 8)),
         country_code = ifelse(species > 10000000000,
                               substr(species, start = 10, stop = 11),
                               substr(species, start = 9, stop = 10))) |>
  mutate(taxon_id = as.numeric(taxon_id),
         bioregion = as.numeric(bioregion),
         country_code = as.numeric(country_code)) |>
  left_join(BR_lookup, multiple = "all") |>
  left_join(spp_lookup, multiple = "all")


##################### abstract statistics ##############################

ic_join <- rep_IC |>
  rename(rep_ic = rep,
         shortfall_ic = shortfall,
         shortfall_perc_ic = shortfall_perc) |>
  dplyr::select(species, rep_ic, shortfall_ic, shortfall_perc_ic)

## carbon percentage change
c_perc <- rep |>
  left_join(ic_join) |>
  filter(spp == "carbon") |>
  filter(rep_ic > 0) |>
  mutate(rep_diff = (rep - rep_ic)/rep_ic) |>
  group_by(spp, scenario, carbon_weight, country_contraint, future) |>
  summarise(rep_diff = rep_diff*100) |>
  arrange(-rep_diff)

## conservation status
b_status <- rep |>
  left_join(ic_join) |>
  filter(species != 999999) |>
  mutate(target_met_ic = ifelse(shortfall_perc_ic < 0, 0, 1),
         target_met = ifelse(shortfall_perc < 0, 0, 1)) |>
  filter(target_met_ic<1) |>
  mutate(improved = (target_met-target_met_ic)) |>
  group_by(spp, scenario, carbon_weight, country_contraint, future) |>
  summarise(improved = sum(improved)/length(unique(species))) |>
  arrange(-improved)
```

