---
title: "Formattable data frame"
author: "Kun Ren"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formattable data frame}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
set.seed(123)
```

Formattable data frames are data frames to be rendered as HTML table with formatter functions applied, which resembles conditional formatting in Microsoft Excel. 

Using `knitr::kable()` or `formattable()`, the data frame can be rendered as HTML table which looks more friendly.

```{r}
library(formattable)
```

```{r}
plain_formatter <- formatter("span")
plain_formatter(c(1, 2, 3))
```

```{r}
width_formatter <- formatter("span",
  style = x ~ style(width = suffix(x, "px")))
width_formatter(c(10, 11, 12))
```
```{r}
sign_formatter <- formatter("span", 
  style = x ~ style(color = ifelse(x > 0, "green", 
    ifelse(x < 0, "red", "black"))))
sign_formatter(c(-1, 0, 1))
```

We can also create another formatter function that makes above-average values bold while leaving others unchanged.

```{r}
above_avg_bold <- formatter("span", 
  style = x ~ style("font-weight" = ifelse(x > mean(x), "bold", NA)))
```






```{r }
library(tidyverse)
rest_table <- read_csv("data/rest_table_results.csv")


t <- rest_table |> 
  mutate(`% total area restored` = round(percent_restored*100, 1),
                     `% deintensification` = round((deintensification/area_restored)*percent_restored*100,1),
                     `% nature` = round((nature/area_restored)*percent_restored*100,1),
                     `% rewetting` = round((rewetting/area_restored)*percent_restored*100,1)) |>
  mutate(country_TF = ifelse(country_TF == "FLEX", "Flexible",
                             ifelse(country_TF == "EVEN", "Even ", "Unconstrained")) ) |>
  mutate(future = ifelse(future == "f455.csv", "Fit for 55", "BAU")) |>
  rename(`burden sharing` = country_TF,
         `restoration scenario` = scenario,
         `production constraints` = future) |>
  dplyr::select(-c(area_restored, nature, rewetting, deintensification, percent_restored)) |>
  #dplyr::select(-c(area_restored)) |> 
  formattable(list(
  `% total area restored` =color_tile("grey","#654321"),
  `% nature` = color_bar("darkgreen"),
  `% rewetting` = color_bar("lightblue"),
  `% deintensification` = color_bar("#FFCC00")))


t
```

## Area formatting

Sometimes, it is useful to apply a formatter function to an area so that all cells in the area share one benchmark. Area formatting is supported through the syntax of `area(row, col) ~ formatter` in the formatter list.

The following example renders the three columns altogether so that they share the same benchmark, not independently.

```{r}
set.seed(123)
df <- data.frame(id = 1:10, 
  a = rnorm(10), b = rnorm(10), c = rnorm(10))
formattable(df, list(area(col = a:c) ~ color_tile("transparent", "pink")))
```

If a one-sided formula is supplied, the function will be applied to all cells.

```{r}
formattable(df[, -1], list(~ percent))
```

# Dynamically generating formatters

Since `formattable()` accepts a list of formatter functions, the list can be dynamically generated. For example, the following code applies row-wise formatting, that is, each row is colored independently.

```{r}
df <- cbind(data.frame(id = 1:10), 
  do.call(cbind, lapply(1:8, function(x) rnorm(10))))
formattable(df, lapply(1:nrow(df), function(row) {
  area(row, col = -1) ~ color_tile("lightpink", "lightblue")
}))
```
